---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# fieldmaps

<!-- badges: start -->
<!-- badges: end -->

`{fieldmaps}` provides an R interface to [fieldmaps](https://fieldmaps.io), a humanitarian geospatial data platform that delivers standardised, edge-matched administrative boundaries for countries worldwide. The package enables easy access to high-quality geospatial data for R users, particularly those working in humanitarian contexts.

## About Fieldmaps

Fieldmaps consolidates authoritative geospatial data from multiple sources into a living, practical and standardised database. Key features include:

- **Edge-matched boundaries**: Globally consistent administrative boundaries that eliminate gaps and overlaps. Uses the OpenStreetMap International ADM0 worldview for edge-matching
- **Two datasets**:
  - [`humanitarian`](https://fieldmaps.io/data/cod): UN OCHA Common Operational Datasets (COD) data prioritizing validated humanitarian datasets. Falls back to geoBoundaries for regions without coverage
  - [`open`](https://fieldmaps.io/data/geoboundaries): geoBoundaries data providing comprehensive global coverage. Suitable for academic or commercial use. 
- **Multiple geometry types**: Polygons, lines, and points for flexible mapping needs

## Installation

You can install the package from [r-universe](https://r-universe.dev/search):

``` r
install.packages('fieldmaps', repos = c('https://epicentre-msf.r-universe.dev', 'https://cloud.r-project.org'))
```

Or the development version from github via the `{pak}` package:

``` r
if (!require("pak", quietly = TRUE)) {
  # binary build of pak from github repository for your system
  install.packages("pak", repos = sprintf("https://r-lib.github.io/p/pak/stable/%s/%s/%s", .Platform$pkgType, R.Version()$os, R.Version()$arch))
}
pak::pak("epicentre-msf/fieldmaps")
```

## Features
 
- **Simple API**: Two main functions for downloading administrative boundaries
- **Flexible country input**: Accepts country names or ISO3 codes (e.g., "Kenya" or "KEN")
- **Multiple admin levels**: Access levels 1-4 depending on data availability
- **Dataset choice**: Select between humanitarian (COD) or open (geoBoundaries) datasets
- **Geometry options**: Download polygons, lines, or points
- **Efficient data access**: Leverages DuckDB to query remote Parquet files without full downloads
- **SF integration**: Returns standard `sf` objects ready for spatial analysis

## Example

### Download a single administrative level

```{r example, fig.width=8, fig.height=6}
library(fieldmaps)

# Download admin level 1 boundaries
adm1 <- get_adm_level("Niger", level = 1)

# features returned by fieldmaps
dplyr::glimpse(adm1)

# plot the sf geometry
plot(adm1$geometry)
```

### Download all available administrative levels

```{r eval=FALSE}
# Download all available admin levels for a country
shps <- get_all_adm_levels("NER") # use country name or ISO3 code

# Access individual levels
shps$ADM1  # Level 1 boundaries
shps$ADM2  # Level 2 boundaries
shps$ADM3  # Level 3 boundaries
```

### Use different datasets and geometry types

```{r eval=FALSE}
# Get boundaries from the open dataset (geoBoundaries)
adm1_open <- get_adm_level("Uganda", level = 1, dataset = "open")

# Get line geometry instead of polygons
adm1_lines <- get_adm_level("Somalia", level = 1, geom = "lines")

# Get point geometry (administrative centers)
adm1_points <- get_adm_level("Ethiopia", level = 1, geom = "points")
```

### Error handling

if you misspell a country name, provide an invalid ISO3 code or request a level that is not available, the package will return an informative error message:

```{r error=TRUE}
get_adm_level("typo")
```

```{r error=TRUE}
get_adm_level("Niger", level = 4)
```

## Technical Details

`{fieldmaps}` uses [DuckDB](https://duckdb.org/) to efficiently query remote Parquet files hosted by fieldmaps without downloading entire datasets. Data is then converted into [`{sf}`](https://r-spatial.github.io/sf/) objects via the [`{geoarrow}`](https://geoarrow.org/geoarrow-r/) R implementation.

This approach minimizes bandwidth requirements and provides fast access to boundaries, making it suitable for use in low-connectivity environments.

## Data Attribution

All data accessed through this package comes from [fieldmaps](https://fieldmaps.io) and includes:

- **Humanitarian dataset**: [UN OCHA Common Operational Datasets (COD)](https://fieldmaps.io/data/cod)
- **Open dataset**: [geoBoundaries data](https://fieldmaps.io/data/geoboundaries)
- Underlying data from OpenStreetMap, national mapping agencies, and other authoritative sources

Please refer to the `src_name`, `src_url`, and `src_lic` fields in the returned data for specific attribution and licensing information for each country.

## Acknowledgements

This R package is a simple interface to the fieldmaps data platform. Many thanks to the fieldmaps maintainer [Max Malynowsky and project contributors](https://fieldmaps.io/about).

## Related R Packages

- [rhdx](https://github.com/dickoa/rhdx) - R client for the Humanitarian Exchange Data platform
- [rgeoboundaries](https://github.com/dickoa/rgeoboundaries) - Access geoBoundaries data directly
- [rnaturalearth](https://github.com/ropensci/rnaturalearth) - Natural Earth map data
- [osmdata](https://github.com/ropensci/osmdata) - OpenStreetMap data access

