---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# fieldmaps

<!-- badges: start -->
<!-- badges: end -->

`{fieldmaps}` provides an R interface to [fieldmaps](https://fieldmaps.io), a humanitarian geospatial data platform that delivers standardized, edge-matched administrative boundaries for countries worldwide. The package enables easy access to high-quality reference data designed for humanitarian and development work, especially in challenging environments with limited connectivity or variable data quality.

## About Fieldmaps

Fieldmaps consolidates authoritative geospatial data from multiple sources into a living, practical database:

- **Data sources**: OpenStreetMap baseline enhanced with UN OCHA Common Operational Datasets (COD), national authoritative sources, and geoBoundaries
- **Edge-matched boundaries**: Globally consistent administrative boundaries that eliminate gaps and overlaps
- **Two datasets**:
  - `humanitarian`: UN OCHA COD data prioritizing validated humanitarian datasets
  - `open`: geoBoundaries data providing comprehensive global coverage
- **Multiple geometry types**: Polygons, lines, and points for flexible mapping needs

The platform specifically targets humanitarian field workers operating under resource constraints, making geographic data accessible even in low-bandwidth environments.

## Installation

You can install the development version of  like so:

``` r
pak::pak("epicentre-msf/")
```

## Features

- **Simple API**: Two main functions for downloading administrative boundaries
- **Flexible country input**: Accepts country names or ISO3 codes (e.g., "Kenya" or "KEN")
- **Multiple admin levels**: Access levels 1-4 depending on data availability
- **Dataset choice**: Select between humanitarian (COD) or open (geoBoundaries) datasets
- **Geometry options**: Download polygons, lines, or points
- **Efficient data access**: Leverages DuckDB to query remote Parquet files without full downloads
- **Automatic validation**: Built-in country code conversion and level checking
- **SF integration**: Returns standard `sf` objects ready for spatial analysis

## Example

### Download a single administrative level

```{r example, fig.width=8, fig.height=6}
library(fieldmaps)

# Download admin level 1 boundaries
adm1 <- get_adm_level("Niger", level = 1)

# features provided by 
dplyr::glimpse(adm1)

# plot the boundaries
plot(adm1$geometry)
```

### Download all available administrative levels

```{r eval=FALSE}
# Download all available admin levels for a country
shps <- get_all_adm_levels("NER") # use country name or ISO3 code

# Access individual levels
shps$ADM1  # Level 1 boundaries
shps$ADM2  # Level 2 boundaries
shps$ADM3  # Level 3 boundaries
```

### Use different datasets and geometry types

```{r eval=FALSE}
# Get boundaries from the open dataset (geoBoundaries)
adm1_open <- get_adm_level("Uganda", level = 1, dataset = "open")

# Get line geometry instead of polygons
adm1_lines <- get_adm_level("Somalia", level = 1, geom = "lines")

# Get point geometry (administrative centers)
adm1_points <- get_adm_level("Ethiopia", level = 1, geom = "points")
```

## Technical Details

**** uses [DuckDB](https://duckdb.org/) to efficiently query remote Parquet files hosted by  without downloading entire datasets. Data is then converted into R `sf` objects efficiently via geoarrow.

This approach minimizes bandwidth requirements and provides fast access to boundaries, making it suitable for use in low-connectivity environments.

## Data Attribution

All data accessed through this package comes from [Fieldmaps](https://.io) and includes:

- **Humanitarian dataset**: UN OCHA Common Operational Datasets (COD)
- **Open dataset**: geoBoundaries data
- Underlying data from OpenStreetMap, national mapping agencies, and other authoritative sources

Please refer to the `src_name`, `src_url`, and `src_lic` fields in the returned data for specific attribution and licensing information for each country.

## Related Packages

- [rgeoboundaries](https://github.com/dickoa/rgeoboundaries) - Access geoBoundaries data directly
- [rnaturalearth](https://github.com/ropensci/rnaturalearth) - Natural Earth map data
- [osmdata](https://github.com/ropensci/osmdata) - OpenStreetMap data access

